Approximation as Differentiation

### 2025-05-05 appendix/notes/auto-diff.tex
Clarify chain rule.

Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$ we can define the following
composite derivatives of $g \comp f: \RR^m \to \RR^k$:

\begin{itemize}
\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x$
\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}$
\end{itemize}

@@

For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$. Regardless of whether the linear maps $\RR^m \linearto \RR^n$ happen to be derivatives, we
can define composition on such maps as:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent and verify that $\tangents$ is indeed a functor. If $f_2(x)$ happens to be the derivative of $f_1$
at $x$ for any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain
rule and $(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.

@@

Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor $\reindex{-}{f}:
\Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{g_y\}_{y \in Y}$ to
$\{g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y \in Y})$
and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
\begin{itemize}
\item $\partial f_x: V_x \to W_{f(x)}$;
\item $\reindex{\partial g}{f}_x: W_{f(x)} \to U_{g(f(x))}$.
\end{itemize}

>>>>

Derivatives respect composition. Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. For any $x \in \RR^m$
we have:

\begin{itemize}
\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
\end{itemize}

@@

For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$. We can define composition on such maps as:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent and verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
\linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and
$(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.

@@

Also recall that $\Fam(f, \FinVect)$ for any function $f: X \to Y$ is the reindexing functor $\reindex{-}{f}:
\Fam(Y, \FinVect) \to \Fam(X, \FinVect)$ sending any family of morphisms $\{\partial g_y\}_{y \in Y}$ to
$\{\partial g_{f(x)}\}_{x \in X}$. Then for any $(f, \partial f): (X, \{V_x\}_{x \in X}) \to (Y, \{W_y\}_{y
\in Y})$ and $(g, \partial g): (Y, \{W_y\}_{y \in Y}) \to (Z, \{U_z\}_{z \in Z})$ we have, for any $x \in X$:
\begin{itemize}
\item $\partial f_x: V_x \to W_{f(x)}$;
\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: W_{f(x)} \to U_{g(f(x))}$.
\end{itemize}
